### 2023-11-12 thesis/sections/2-background.tex
background

\subsection{Referring expressions}
\cmtDK[inline]{what are referring expressions and how are they studied (generation, understanding, ...)}
\cmtDK[inline]{incremental GRE algorithm from Dale}

>>>>

\subsection{Referring expressions}
\label{sec:background:re}
A specific case of grounding linguistic information in visual input are referring expressions.
While a general linguistic expression can be grounded in multiple regions of a visual scene, referring expressions are expected to refer to a unique region \citep{Sanchez2022}.
Given for example a visual scene with multiple pens lying on a table, the word 'pen' can be grounded in all visible objects.
A referring expression like 'the red pen on the left' in contrast would only refer to one specific object.
Referring expressions are however not limited to refer to one single entity, but might also refer to a group of objects as for instance with 'the green pens'.
Still they refer to a specific subset of entities and concepts.

Hereby, \citet{Krahmer2012} classify referring expressions into four categories:
\emph{One-place predicates} refer to one single entity in the visual scene without.
By this, the entity is identifiable without using another entity as basis.
Often, the referring expressions includes intrinsic attributes of an object or absolute locations in the scene as in 'the red pen on the left'.
\emph{Relational} referring expressions refer to entities via other entities in the scene.
Accordingly, a relating between the target entity and distractors is described.
This includes for example spatial relations like 'the pen behind the cup' or relations based on the entities' properties like 'the largest pen'.
Referring expressions can also identify \emph{sets} of entities.
Instead of uniquely referring to on single pen, the referring expression 'the green pens' refers to all objects that are pens and that are green.
Finally, \emph{gradable} referring expressions involve gradable properties as opposed to absolute values.
For example, given the referring expression 'the tall building', 'tall' is a gradable property because it can be measured on a scale.
The success of the referring expression depends on how well it captures the intended level of height compared to other buildings.
By this, vagueness is introduced and referring expressions depend on context, either present in the visual scene or pragmatic context.

One of the main challenges of referring expressions lies in ambiguity.
Ambiguity is introduced in two forms.
First, there is an inherent difference in natural language and the real world it is grounded it, also described in the symbol grounding problem.
The nature of the world is continuous, while natural language being a formal symbol system uses discrete symbols.
In especially, natural language utilizes a very limited set symbols to represent the world.
When trying to map continuous concepts to this limited set of discrete symbols, information nesecarilly gets lost.
Apparent examples are the color systems in human languages.
While color grades exist on an infinite spectrum in the real world, languages combine certain color grades into categories \citep{Zaslavsky2018}.
The word 'red' for example refers to many different shades of red.
Even more, the shades of color it actually refers to depends also on the pragmatic context and the people uttering it \citep{Monroe2017}.
The same also applies to spatial references.
References like 'near to' or 'left to' might refer to objects that are touching each other, but in other contexts also are kilometers away from each other.

The other form of ambiguity is introduced through under-specification of the speakers (e.g. \citet{Dobnik2021}).
When interlocutors interact in a dialogue, they share some common knowledge.
The first interlocutor might presuppose that the second interlocutor knows a certain fact.
Based on this, they might utter a referring expression that together with this shared fact uniquely identifies an entity.
However, the referring expression on its own is under-specified and might refer to multiple entities.
Given for example the following situation:
Two people are in a room with their beverages standing on the table and Person A requests his beverage from Person B.
A referring expression like 'the beverage close to me' would uniquely identify on of the beverages.
Despite that, Person A might more likely use 'the beverage' to refer to theirs.
On its own, the referring expression is under-specified and can refer to both of the present beverages.
However, both persons share the fact that the targeted beverage belongs to person A, which disambiguates the referring expression in this context.
In another context, the shared fact would be that Person A wants to try the beverage of Person B.
Again, this fact would disambiguate the referring expression, but now refer to the second beverage.
The referring expression is the same in both contexts, but refers to two different entities.
This creates a challenge for computational models, that need to understand and generate referring expressions if they are only presented with the text, but not with the shared context.

The research of referring expressions can be split into two fields: referring expression generation (REG) and referring expression comprehension (REC).
In \textbf{referring expression generation}, computational models are trained to generate a referring expression that uniquely identifies an entity given some perceptual, often visual input.
The research goes back into the 70s, where \citet{Winograd1972} developed an algorithm to generate referring expressions step by step, taking into account the context and the information available at each stage of the generation process.
A similar approach is described in \citep{Dale1995}.
Their incremental algorithm is based on the salience of the properties of the present target object and distractor objects.
The most salient properties are considered first, and added to the referring expression until the referring expression is unambiguous.
In other words, the object is described unambiguously using the lowest number of words.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\linewidth]{figures/CLEVR_dale-5.png}
    \caption{The target object in this scene is the \emph{small turquoise sphere}. Using the incremental GRE-algorithm, the target object can be uniquely referred to as \emph{turquoise sphere}}
    \label{fig:incremental-algorithm}
\end{figure}

Given for example the image in Figure \ref{fig:incremental-algorithm}.
Objects in this scene have three attributes that we consider in the following order of salience: a shape, a color and a size.
This order defines, which attributes can be left out, while still identifying the object uniquely.
The target object in the given scene is the \emph{small turquoise sphere}.
In the first step, the attribute with the highest salience, the shape, is selected from the target object.
This produces the referring expression \emph{sphere}.
In the next step, it is checked if there are also distractors, described by this referring expression.
In our case, both cubes and the cylinder are already disambiguated, but the purple sphere can still be described by the produced referring expression.
Therefore, the algorithm also adds the next lower salient attribute, the color to the referring expression, which produces the \emph{turquoise sphere}.
With this, also the last remaining distractor is as well disambiguated, and the referring expression describes the target object uniquely.
The size, the least salient property is therefore not necessary in the referring expression.
This algorithm will be utilized in this thesis and will be referred to as \emph{the incremental GRE-algorithm}.

Additional to generating an efficient disambiguating referring expressions, a challenge also lies in processing the visual input.
This especially applies to spatial and geometric positions and relations of objects in a scene, where models struggle to extract geometric information from visual features \citep{Kelleher2017}.
